Appendix 3: Example of a spatially varying effect of climate on citizen science count data with sdmTMB

### 2022-01-08 doc/appendix-snow-sdmTMB.Rmd
add maps of upper and lower CI for spatially varying effect

```{r packages, message=FALSE, warning=FALSE, cache=FALSE}
library(ggplot2)
library(dplyr)
library(sdmTMB)
library(sf)
```

@@

We can then compress the annual predictions into mean values for plotting the spatially varying effect of NAO.
First, we extract the overall coefficients using the `b <- tidy(m)` and add them together with the spatially varying coefficients `zeta_s`.

```{r p-mean, message=FALSE, echo=TRUE, eval=TRUE}
b <- tidy(m)
p_mean <- p %>%
  group_by(CBCID) %>%
  summarise(
    X = mean(X), Y = mean(Y),
    mean_est_count = exp(mean(est)),
    nao_effect = exp(b[b$term == "nao", 2] + mean(zeta_s))
  )
```

```{r plot-map, echo=TRUE, eval=TRUE}
ggplot(data = p_proj) +
  geom_point(
    data = p_mean,
    aes(X, Y, colour = nao_effect, size = mean_est_count), alpha = 0.5
  ) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000), # adjusts space on sides
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(guide = guide_colourbar(
    direction = "horizontal", title.position = "top"
  )) +
  guides(size = "none") +
  labs(colour = "NAO effect\non Snowy Owl count") +
  theme(legend.position = c(0.1, 0.1), axis.title = element_blank())
```

>>>>

```{r packages, message=FALSE, warning=FALSE, cache=FALSE}
library(ggplot2)
library(patchwork)
library(dplyr)
library(sdmTMB)
library(sf)
```

@@

We can then compress the annual predictions into mean values for plotting the spatially varying effect of NAO.
Most simply, we can extract the overall coefficients using the `b <- tidy(m)` and add them together with the spatially varying coefficients `zeta_s`.
However, this method does not give us spatially explicit estimates of uncertainty.
To do this, we can use the output from the joint precision matrix simulation draws by using the `predict()` function again but this time indicating a number of simulation draws (e.g., `sims = 500`) and what component we want estimated (`sims_var = "zeta_s"`). 
We also extract the same number of drams for (this time `n_sims = 500`) in the `spread_sims()` function for the main effect of NAO. 
The simulation draws for both should then be combined before adding the median and lower and upper quantiles from these draws to our prediction data frame.

```{r p-mean, message=FALSE, echo=TRUE, eval=TRUE}
b <- tidy(m)

n_sims <- 500L
zeta_sim <- predict(m, sims = n_sims, sims_var = "zeta_s")
sims <- spread_sims(m, n_sims = n_sims)
beta <- sims$nao
combined <- beta + t(zeta_sim)

p$zeta_sim <- apply(t(zeta_sim), 2, median)
p$nao_effect_sim <- apply(combined, 2, median)
p$nao_effect_lwr <- apply(combined, 2, quantile, probs = 0.10)
p$nao_effect_upr <- apply(combined, 2, quantile, probs = 0.90)

p_mean <- p %>%
  group_by(CBCID) %>% summarise(
    year = max(year),
    X = mean(X), Y = mean(Y),
    mean_est_count = exp(mean(est)),
    nao_effect = b[b$term == "nao", 2] + mean(zeta_s),
    nao_sim = mean(nao_effect_sim),
    nao_lwr = mean(nao_effect_lwr),
    nao_upr = mean(nao_effect_upr)
  )
```

```{r plot-map, echo=TRUE, eval=TRUE, out.width="100%"}
ggplot(data = p_proj) +
  geom_point(
    data = p_mean,
    aes(X, Y, colour = nao_effect, size = mean_est_count), alpha = 0.5
  ) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000), # adjusts space on sides
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(
    limit = c(min(p_mean$nao_lwr), max(p_mean$nao_upr)), 
    guide = guide_colourbar(direction = "horizontal", title.position = "top")
  ) +
  guides(size = "none") +
  labs(colour = "Estimated effect in log space") +
  ggtitle("Map of combined main and spatially varying effect of NAO on Snowy Owl count") +
  theme(legend.position = c(0.1, 0.07), axis.title = element_blank())
```

```{r plot-map-ci, echo=F, eval=TRUE, out.width="100%"}
p1 <- ggplot(data = p_proj) +
  geom_point(
    data = p_mean,
    aes(X, Y, colour = nao_lwr, size = mean_est_count), alpha = 0.5
  ) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000), # adjusts space on sides
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(
    limit = c(min(p_mean$nao_lwr), max(p_mean$nao_upr)), 
    guide = guide_colourbar(direction = "horizontal", title.position = "top")
  ) +
  guides(size = "none", colour = "none") +
  # labs(colour = "Lower 90% CI") +
  ggtitle("Lower 80% CI") +
  theme(legend.position = c(0.1, 0.1), axis.title = element_blank())

p2 <- ggplot(data = p_proj) +
  geom_point(
    data = p_mean,
    aes(X, Y, colour = nao_upr, size = mean_est_count), alpha = 0.5
  ) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000), # adjusts space on sides
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(
    limit = c(min(p_mean$nao_lwr), max(p_mean$nao_upr)), 
    guide = guide_colourbar(direction = "horizontal", title.position = "top")
  ) +
  guides(size = "none", colour = "none") +
  # labs(colour = "Upper 90% CI") +
  ggtitle("Upper 80% CI") +
  theme(legend.position = c(0.1, 0.1), axis.title = element_blank())

p1 + p2 + patchwork::plot_layout()
```
